---
title: "Plot data preparation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plot-data-preparation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
# Setting eval=FALSE for all chunks to prevent large file downloads during package checks
knitr::opts_chunk$set(eval = FALSE)
#library(Plot2Map)
devtools::load_all()
```


## Introduction

In this page we showcase the different user data formats of plot estimates of AGB, here refereed as plot data, possible to be used in the Plot2Map workflow.

## Plot data formats

### Point data formatted dataframe

In this case there is already an AGB estimate from the user. This format is exemplified with the sample plots dataframe available with the package accessible with:

```{r plots}
head(plots) 

```

This format correspond to a dataframe containing sample plots with an unique plot ID, longitude and latitude point data, an estimated Above Ground Biomass in t/ha, average year of inventory, and plot size in ha.



### Point data unformatted dataframe



In this case there is also a dataframe in a similar format to the previous case, but with unformatted data over which the user will be asked about the specific column index of required plot variables such as unique plot ID, longitude, latitude, AGB of the plot, plot size, inventory year. This can be exemplified below:

```{r unformatted plots1, eval=FALSE}

unformatted_plots <- utils::read.csv(sample_file("SampleUnformattedPlots.csv"))
head(unformatted_plots)
#      Provincia                          Estrato.Florestal id_parcela Ycoordinate Xcoordinate Vt..m.3.ha.
# 1 Cabo Delgado Floresta (semi-) decidua, incluindo Miombo  CD0772003   -10.60656    40.53664       39.90
# 2 Cabo Delgado Floresta (semi-) decidua, incluindo Miombo  CD0772002   -10.60657    40.53572       24.65
# 3 Cabo Delgado Floresta (semi-) decidua, incluindo Miombo  CD0770543   -10.60666    40.52744        4.09
# 4 Cabo Delgado Floresta (semi-) decidua, incluindo Miombo  CD0770542   -10.60667    40.52652       20.06
# 5 Cabo Delgado Floresta (semi-) decidua, incluindo Miombo  CD0772004   -10.60747    40.53664        0.00
# 6 Cabo Delgado Floresta (semi-) decidua, incluindo Miombo  CD0772001   -10.60747    40.53573       64.15
#   Vc..m.3.ha. AGB..ton.ha. BGB..ton.ha. year size
# 1        2.53        32.61        21.73 2016  0.1
# 2        0.77        21.44        13.32 2016  0.1
# 3        0.00         4.41         2.96 2016  0.1
# 4        2.38        17.77        10.56 2016  0.1
# 5        0.00         0.00         0.00 2016  0.1
# 6        0.97        77.74        27.38 2016  0.1

```

The following step will request the user to select the requested columns:

```{r unformatted plots2, eval=FALSE}

formatted_plots <- RawPlots(unformatted_plots)
# Which column is your unique Plot ID? 
# 
#  1: Manual entry        2: Provincia           3: Estrato.Florestal
#  4: id_parcela          5: Ycoordinate         6: Xcoordinate      
#  7: Vt..m.3.ha.         8: Vc..m.3.ha.         9: AGB..ton.ha.     
# 10: BGB..ton.ha.       11: year               12: size             
# 
# 
# Selection: 4
# Which column is your plot AGB? 
# 
#  1: Manual entry        2: Provincia           3: Estrato.Florestal
#  4: id_parcela          5: Ycoordinate         6: Xcoordinate      
#  7: Vt..m.3.ha.         8: Vc..m.3.ha.         9: AGB..ton.ha.     
# 10: BGB..ton.ha.       11: year               12: size             
# 
# 
# Selection: 9
# Select longitude column 
# 
#  1: Manual entry        2: Provincia           3: Estrato.Florestal
#  4: id_parcela          5: Ycoordinate         6: Xcoordinate      
#  7: Vt..m.3.ha.         8: Vc..m.3.ha.         9: AGB..ton.ha.     
# 10: BGB..ton.ha.       11: year               12: size             
# 
# 
# Selection: 6
# Which column is your latitude? 
# 
#  1: Manual entry        2: Provincia           3: Estrato.Florestal
#  4: id_parcela          5: Ycoordinate         6: Xcoordinate      
#  7: Vt..m.3.ha.         8: Vc..m.3.ha.         9: AGB..ton.ha.     
# 10: BGB..ton.ha.       11: year               12: size             
# 
# 
# Selection: 5
# Select plot size column 
# 
#  1: Manual entry        2: Provincia           3: Estrato.Florestal
#  4: id_parcela          5: Ycoordinate         6: Xcoordinate      
#  7: Vt..m.3.ha.         8: Vc..m.3.ha.         9: AGB..ton.ha.     
# 10: BGB..ton.ha.       11: year               12: size             
# 
# 
# Selection: 12
# Select year column 
# 
#  1: Manual entry        2: Provincia           3: Estrato.Florestal
#  4: id_parcela          5: Ycoordinate         6: Xcoordinate      
#  7: Vt..m.3.ha.         8: Vc..m.3.ha.         9: AGB..ton.ha.     
# 10: BGB..ton.ha.       11: year               12: size             
# 
# 
# Selection: 11


```

Note how after user input we get a similar data format to the previous section:

```{r unformatted plots3, eval=FALSE}

tail(formatted_plots)
#      PLOT_ID  POINT_X   POINT_Y AGB_T_HA SIZE_HA FEZ GEZ AVG_YEAR
# 3267      21 32.35752 -26.54318    54.13     0.1  NA  NA     2016
# 3268      24 32.35853 -26.54318    13.90     0.1  NA  NA     2016
# 3269      12 32.75884 -26.65187    92.66     0.1  NA  NA     2016
# 3270      13 32.75984 -26.65187   112.87     0.1  NA  NA     2016
# 3271      11 32.75884 -26.65277    56.08     0.1  NA  NA     2016
# 3272      14 32.75984 -26.65277    65.59     0.1  NA  NA     2016

```


This format correspond to a dataframe containing sample plots with an ID, longitude and latitude point data, an estimated Above Ground Biomass in t/ha, average year of plot survey, and plot size in ha.



### Plot corner coordinates

This is the plot data format found in [Labriere et al. 2018](https://doi.org/10.1109/JSTARS.2018.2851606) sample data. It contains the corner coordinates of a bounding box defining each plot:


```{r polygon_corner_coords, eval=FALSE}

plots_corners <- utils::read.csv(sample_file("PolyTropiSAR.csv")) # Labriere et al. 2018 plots sample data
head(plots_corners)
#     id  POINT_X  POINT_Y  X
# 1 NOU01 312514.3 449724.3 NA
# 2 NOU01 312517.6 449824.2 NA
# 3 NOU02 313704.3 451320.5 NA
# 4 NOU02 313752.2 451408.3 NA
# 5 NOU02 313800.0 451496.1 NA
# 6 NOU02 313847.9 451583.9 NA

plots_AGB <-  utils::read.csv(sample_file("PolyTropiAGB.csv")) # plot-level AGB
head(plots_AGB)
#   PLOT_ID AVG_YEAR AGB_T_HA
# 1   NOU01     2010   431.90
# 2   NOU02     2010   368.41
# 3   NOU03     2010   525.65
# 4   NOU04     2010   429.87
# 5   NOU05     2010   234.60
# 6   NOU06     2010   288.10

SRS <- 32622 # plot data source spatial reference system
# Converting coordinates from source SRS (32622) to EPSG:4326 / WGS84
plots_corners_AGB$AGB_T_HA <- plots_AGB$AGB_T_HA
plots_corners_AGB$AVG_YEAR <- plots_AGB$AVG_YEAR
 
head(plots_corners_AGB)
#       ID PLOT_ID         SIZE_HA   POINT_X  POINT_Y AGB_T_HA AVG_YEAR
# NOU01  1   NOU01  2.0000014 [ha] -52.68839 4.067844   431.90     2010
# NOU02  1   NOU02 10.0000058 [ha] -52.67643 4.085602   368.41     2010
# NOU03  1   NOU03  6.0000008 [ha] -52.68235 4.039235   525.65     2010
# NOU04  1   NOU04 12.0000037 [ha] -52.68352 4.082891   429.87     2010
# NOU05  1   NOU05  0.2499999 [ha] -52.67710 4.041068   234.60     2010
# NOU06  1   NOU06  0.2499998 [ha] -52.67789 4.039038   288.10     2010

```



### Plot data contains tree-level measurements

This format of plot data contains tree-level allometric measurements that will be used to 
estimate mean AGB and standard deviation by plot ID, using [BIOMASS](https://umr-amap.github.io/BIOMASS/) package's height-diameter model:


```{r plot_trees1, eval=FALSE}

# Formatted data example
plotsTree <- utils::read.csv(sample_file("SampleTree.csv"))
head(plotsTree)
#     id      genus    species  diameter  size                 fez      gez year
# 1 BSP1 Terminalia  bellirica  3.501409 10000 tropical rainforest tropical 1996
# 2 BSP1   Ziziphus   oenoplia  3.819719 10000 tropical rainforest tropical 1996
# 3 BSP1    Aporosa lindleyana 16.870424 10000 tropical rainforest tropical 1996
# 4 BSP1      Ixora  brachiata  4.138029 10000 tropical rainforest tropical 1996
# 5 BSP1   Wrightia    arborea  5.411268 10000 tropical rainforest tropical 1996
# 6 BSP1      Ixora  brachiata  3.183099 10000 tropical rainforest tropical 1996

xyTree <- utils::read.csv(sample_file("SampleTreeXY.csv"))
head(xyTree)
#     id        y        x
# 1 BSP1 14.36806 74.91944
# 2 BSP1 14.36806 74.91944
# 3 BSP1 14.36806 74.91944
# 4 BSP1 14.36806 74.91944
# 5 BSP1 14.36806 74.91944
# 6 BSP1 14.36806 74.91944

# Unformatted data example
TreeRaw <- utils::read.csv(sample_file("KarnatakaForest.csv"))
head(TreeRaw)
#   plotId     treeId     family       genus        species     d_cm      lat     long size_m2 year
# 1   BSP1   BSP1_248 Clusiaceae    Garcinia         indica 3.183099 14.36806 74.91944   10000 1998
# 2   BSP1   BSP1_269  Rubiaceae       Ixora      brachiata 3.183099 14.36806 74.91944   10000 1998
# 3   BSP1 BSP1_303_B  Rubiaceae       Meyna      laxiflora 3.183099 14.36806 74.91944   10000 1998
# 4   BSP1    BSP1_29  Lauraceae Alseodaphne semecarpifolia 3.501409 14.36806 74.91944   10000 1998
# 5   BSP1    BSP1_30  Lauraceae Alseodaphne semecarpifolia 3.501409 14.36806 74.91944   10000 1998
# 6   BSP1    BSP1_31  Lauraceae Alseodaphne semecarpifolia 3.501409 14.36806 74.91944   10000 1998

rawTree <- RawPlotsTree(TreeRaw) # User is be asked to select the requested columns

plotTree <- rawTree[[1]]
head(plotTree)
#   id       genus        species diameter  size fez gez year
# 1 11    Garcinia         indica 3.183099 10000  NA  NA 1998
# 2 11       Ixora      brachiata 3.183099 10000  NA  NA 1998
# 3 11       Meyna      laxiflora 3.183099 10000  NA  NA 1998
# 4 11 Alseodaphne semecarpifolia 3.501409 10000  NA  NA 1998
# 5 11 Alseodaphne semecarpifolia 3.501409 10000  NA  NA 1998
# 6 11 Alseodaphne semecarpifolia 3.501409 10000  NA  NA 1998

xyTree <- rawTree[[2]]
head(xyTree)
#   id        x        y
# 1 11 74.91944 14.36806
# 2 11 74.91944 14.36806
# 3 11 74.91944 14.36806
# 4 11 74.91944 14.36806
# 5 11 74.91944 14.36806
# 6 11 74.91944 14.36806

# Calculate mean AGB and SD by PLOT_ID using BIOMASS package's height-diameter model
plots_trees_agb_sd <- sd_tree(plotTree, xyTree, "India")
# The reference dataset contains 289 wood density values
# Your taxonomic table contains 122 taxa
# No tree height data found in original plot data. Calculating height using BIOMASS height-diameter model.
# Warning message:
# In BIOMASS::getWoodDensity(genus = plot$genus, species = plot$species,  :
#   DRYAD data only stored 289 wood density values in your region of interest. You could provide additional wood densities (parameter addWoodDensityData) or widen your region (region="World")

print(plots_trees_agb_sd)
#    PLOT_ID  POINT_X  POINT_Y SIZE_HA AVG_YEAR   AGB_T_HA    sdTree
# 1        1 74.91944 14.36806       1     1998 185.527727 12.248335
# 2        2 74.62778 15.16667       1     1998 301.968606 19.298196
# 3        3 74.63667 14.24944       1     1998 160.733560 17.332449
# 4        4 74.66806 14.24528       1     1998 351.992736 20.154294
# 5        5 75.11806 14.28333       1     1998 305.638906 20.040947
# 6        6 75.13583 14.28389       1     1998 425.775593 38.229345
# 7        7 74.87222 14.42500       1     1998 234.641786 13.436895
# 8        8 75.30806 14.44583       1     1998   5.398583  1.654017
# 9        9 75.17222 14.34861       1     1998  81.813532  9.714730
# 10      10 75.28639 14.08750       1     1998  73.081759  6.591937

```



### Plot data contains tree-level measurements and nested plots (sub-plots)

This format of plot data is similar to the previous one but contains nested plot (sub-plots) data. Let's exemplify it by using the following sample shapefile: 

```{r plot_trees2, eval=FALSE}

cent <- sf::st_read(sample_file("SampleCentroid.shp")) # Sub-plot centroids

print(cent)
# Simple feature collection with 5 features and 23 fields
# Geometry type: POINT
# Dimension:     XY
# Bounding box:  xmin: 272004.1 ymin: 350052.3 xmax: 272092.6 ymax: 350087.1
# Projected CRS: OSGB36 / British National Grid
#                               POINT_GUID     GRID_REF POINT_NUMB VISIT_STAT REASON_FOR PEG_LEFT
# 1 {89B25687-EDA0-4704-B00D-93EDDA162779} sh7202950089          1          3          1        1
# 2 {22E69595-D9BE-4BB1-8660-E2E7457E2068} sh7201850062          1          3          1        1
# 3 {B4980FC2-4DA9-4897-93E2-07D03233D2F4} sh7209250061          1          3          1        1
# 4 {224FD85B-6B64-4075-B51F-BA2D390AB4EF} sh7208750054          1          3          1        1
# 5 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} sh7200750088          1          3          1        1
#   PEG_NOT_LE                                  PEG_DESCRI                              PLOT_GUID OBJECTID
# 1          0                   NW of v. slight open area {ABC3BE55-35CC-430B-A9C2-5BA9437286B6}        1
# 2          0                   5m N of edge of mossy gap {E76DFF7B-39B4-45A5-85B8-F7BC38A429D5}        2
# 3          0      W of large SS in shallow plough furrow {6EFDF9E3-F49D-4ED1-8BEB-F53D2A896BBA}        3
# 4          0 in plough ditch with slight gap in row to E {47011DCB-7ADF-4C3D-B621-74AA98265CA4}        4
# 5          0                      N from small mossy gap {EE0EF576-D6FF-4603-864F-23DC93C6A74D}        5
#   INACCESSIB ACCESS_STA GENERATE_N HEIGHT TREE_COUNT ESTIMATED_ TREE_COU_1 RECORD_STA FIELDS_STA
# 1          0          1          0      0          0          0          0        new       <NA>
# 2          0          1          0      0          0          0          0        new       <NA>
# 3          0          1          0      0          0          0          0        new       <NA>
# 4          0          1          0      0          0          0          0        new       <NA>
# 5          0          1          0      0          0          0          0        new       <NA>
#   RESURVEY_P FROZEN POINT_STUM ACCESS_COM                  geometry
# 1          0      0         21       <NA> POINT (272028.2 350085.6)
# 2          0      0         49       <NA> POINT (272013.6 350060.7)
# 3          0      0          8       <NA> POINT (272092.6 350063.9)
# 4          0      0         10       <NA>   POINT (272086 350052.3)
# 5          0      0         22       <NA> POINT (272004.1 350087.1)

tree <- read.csv(sample_file("SampleTreeNested.csv")) # Tree data per sub-plot
str(tree)
# 'data.frame':	153 obs. of  50 variables:
#  $ TREE_GUID..              : chr  "{90106646-EC43-4E04-B3DB-F4DBE0F37FD3}" "{3B18375F-B224-420E-92D0-0D7A0AC51B4A}" "{BCADC7ED-3C2B-41E1-9118-D3A74AAE8353}" "{DE39854B-074C-4CAE-B32C-42AC66063D68}" ...
#  $ TREE_TYPE                : chr  "1" "1" "1" "1" ...
#  $ SPIS                     : chr  "SS" "SS" "SS" "SS" ...
#  $ STRY                     : chr  "4" "4" "4" "4" ...
#  $ TREE_ALIVE               : chr  "1" "1" "1" "1" ...
#  $ DEAD_TREE_CAUSE          : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ GROUP_TYPE               : chr  "1" "1" "1" "1" ...
#  $ CONIFER_STEM_STRAIGHTNESS: chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ LOWEST_DEAD_BRANCH_HEIGHT: chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ POINT_GUID..             : chr  "{197A519B-7AF1-4374-BDD8-81FA03CCC54F}" "{197A519B-7AF1-4374-BDD8-81FA03CCC54F}" "{197A519B-7AF1-4374-BDD8-81FA03CCC54F}" "{197A519B-7AF1-4374-BDD8-81FA03CCC54F}" ...
#  $ TREE_OR_STUMP            : int  1 1 1 1 1 1 1 1 1 1 ...
#  $ DECAY_CLASS              : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ STUMP_DECAY_CLASS        : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ COPPICE_STOOL            : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ LINKED_TREE_GUID..       : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ TREE_LOCATION            : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ OBJECTID..               : int  34 24 25 26 27 28 29 30 31 23 ...
#  $ OUTSIDE_PLOT             : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ HEIGHT                   : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ TREE_HEIGHT              : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ UC_HEIGHT                : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ LC_HEIGHT                : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ SHAPE..                  : chr  "Point" "Point" "Point" "Point" ...
#  $ DBH                      : chr  "11" "11" "10" "12" ...
#  $ DIAMETER_ONE             : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ DIAMETER_TWO             : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ SPECIES_GROUP            : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ EXCESSIVE_LEAN           : chr  "0" "0" "0" "0" ...
#  $ WINDSNAPPED              : chr  "0" "0" "0" "0" ...
#  $ COPPICE_STOOL_GUID..     : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ CROWN_DIAMETER_ONE       : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ CROWN_DIAMETER_TWO       : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ RECORD_STATUS            : chr  "new" "new" "new" "new" ...
#  $ FIELDS_STATUS            : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ TIMBER_HEIGHT            : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ TREE_RESURVEY            : int  1 1 1 1 1 1 1 1 1 1 ...
#  $ GOOD_FELLING_PRACTICE    : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ COPPICE_STOOL_DIAMETER   : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ RANGE_DBH                : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ RANGE_HEIGHT             : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ RANGE_CROWN_DIAMETER_1   : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ RANGE_CROWN_DIAMETER_2   : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ RANGE_DIAMETER_ONE       : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ RANGE_STUMP_HEIGHT       : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ STUMP_TYPE               : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ RANGE_DIAMETER_TWO       : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ SIZE_RANK                : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ PERCENTAGE_HEIGHT_REM    : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ STUMP_AGE                : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...
#  $ IS_STUMP_AGE             : chr  "<Null>" "<Null>" "<Null>" "<Null>" ...

TreeData <- Nested(cent, tree)
# which column is your unique Plot ID? 
# 
#  1: TREE_GUID..                 2: TREE_TYPE                   3: SPIS                     
#  4: STRY                        5: TREE_ALIVE                  6: DEAD_TREE_CAUSE          
#  7: GROUP_TYPE                  8: CONIFER_STEM_STRAIGHTNESS   9: LOWEST_DEAD_BRANCH_HEIGHT
# 10: POINT_GUID..               11: TREE_OR_STUMP              12: DECAY_CLASS              
# 13: STUMP_DECAY_CLASS          14: COPPICE_STOOL              15: LINKED_TREE_GUID..       
# 16: TREE_LOCATION              17: OBJECTID..                 18: OUTSIDE_PLOT             
# 19: HEIGHT                     20: TREE_HEIGHT                21: UC_HEIGHT                
# 22: LC_HEIGHT                  23: SHAPE..                    24: DBH                      
# 25: DIAMETER_ONE               26: DIAMETER_TWO               27: SPECIES_GROUP            
# 28: EXCESSIVE_LEAN             29: WINDSNAPPED                30: COPPICE_STOOL_GUID..     
# 31: CROWN_DIAMETER_ONE         32: CROWN_DIAMETER_TWO         33: RECORD_STATUS            
# 34: FIELDS_STATUS              35: TIMBER_HEIGHT              36: TREE_RESURVEY            
# 37: GOOD_FELLING_PRACTICE      38: COPPICE_STOOL_DIAMETER     39: RANGE_DBH                
# 40: RANGE_HEIGHT               41: RANGE_CROWN_DIAMETER_1     42: RANGE_CROWN_DIAMETER_2   
# 43: RANGE_DIAMETER_ONE         44: RANGE_STUMP_HEIGHT         45: STUMP_TYPE               
# 46: RANGE_DIAMETER_TWO         47: SIZE_RANK                  48: PERCENTAGE_HEIGHT_REM    
# 49: STUMP_AGE                  50: IS_STUMP_AGE               
# 
# Selection: 10
# which column is your unique  DBH (cm)? 
# 
#  1: TREE_GUID..                 2: TREE_TYPE                   3: SPIS                     
#  4: STRY                        5: TREE_ALIVE                  6: DEAD_TREE_CAUSE          
#  7: GROUP_TYPE                  8: CONIFER_STEM_STRAIGHTNESS   9: LOWEST_DEAD_BRANCH_HEIGHT
# 10: POINT_GUID..               11: TREE_OR_STUMP   plots <- MeasurementErr(plotTree, xyTree, 'Europe')           12: DECAY_CLASS              
# 13: STUMP_DECAY_CLASS          14: COPPICE_STOOL              15: LINKED_TREE_GUID..       
# 16: TREE_LOCATION              17: OBJECTID..                 18: OUTSIDE_PLOT             
# 19: HEIGHT                     20: TREE_HEIGHT                21: UC_HEIGHT                
# 22: LC_HEIGHT                  23: SHAPE..                    24: DBH                      
# 25: DIAMETER_ONE               26: DIAMETER_TWO               27: SPECIES_GROUP            
# 28: EXCESSIVE_LEAN             29: WINDSNAPPED                30: COPPICE_STOOL_GUID..     
# 31: CROWN_DIAMETER_ONE         32: CROWN_DIAMETER_TWO         33: RECORD_STATUS            
# 34: FIELDS_STATUS              35: TIMBER_HEIGHT              36: TREE_RESURVEY            
# 37: GOOD_FELLING_PRACTICE      38: COPPICE_STOOL_DIAMETER     39: RANGE_DBH                
# 40: RANGE_HEIGHT               41: RANGE_CROWN_DIAMETER_1     42: RANGE_CROWN_DIAMETER_2   
# 43: RANGE_DIAMETER_ONE         44: RANGE_STUMP_HEIGHT         45: STUMP_TYPE               
# 46: RANGE_DIAMETER_TWO         47: SIZE_RANK                  48: PERCENTAGE_HEIGHT_REM    
# 49: STUMP_AGE                  50: IS_STUMP_AGE               
# 
# Selection: 24
# which column is your unique Tree Height (m)? 
# 
#  1: TREE_GUID..                 2: TREE_TYPE                   3: SPIS                     
#  4: STRY                        5: TREE_ALIVE                  6: DEAD_TREE_CAUSE          
#  7: GROUP_TYPE                  8: CONIFER_STEM_STRAIGHTNESS   9: LOWEST_DEAD_BRANCH_HEIGHT
# 10: POINT_GUID..               11: TREE_OR_STUMP              12: DECAY_CLASS              
# 13: STUMP_DECAY_CLASS          14: COPPICE_STOOL              15: LINKED_TREE_GUID..       
# 16: TREE_LOCATION              17: OBJECTID..                 18: OUTSIDE_PLOT             
# 19: HEIGHT                     20: TREE_HEIGHT                21: UC_HEIGHT                
# 22: LC_HEIGHT                  23: SHAPE..                    24: DBH                      
# 25: DIAMETER_ONE               26: DIAMETER_TWO               27: SPECIES_GROUP            
# 28: EXCESSIVE_LEAN             29: WINDSNAPPED                30: COPPICE_STOOL_GUID..     
# 31: CROWN_DIAMETER_ONE         32: CROWN_DIAMETER_TWO         33: RECORD_STATUS            
# 34: FIELDS_STATUS              35: TIMBER_HEIGHT              36: TREE_RESURVEY            
# 37: GOOD_FELLING_PRACTICE      38: COPPICE_STOOL_DIAMETER     39: RANGE_DBH                
# 40: RANGE_HEIGHT               41: RANGE_CROWN_DIAMETER_1     42: RANGE_CROWN_DIAMETER_2   
# 43: RANGE_DIAMETER_ONE         44: RANGE_STUMP_HEIGHT         45: STUMP_TYPE               
# 46: RANGE_DIAMETER_TWO         47: SIZE_RANK                  48: PERCENTAGE_HEIGHT_REM    
# 49: STUMP_AGE                  50: IS_STUMP_AGE               
# 
# Selection: 20
# Warning: NAs introduced by coercion
# Enter tree genus: Picea
# Enter tree species: sitchensis
# Enter plot size in m2: 100

plotTree <- TreeData[[1]]
head(plotTree)
#                                       id genus    species diameter size fez gez year   height
# 1 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} Picea sitchensis       11  100  NA  NA 2010 11.96667
# 2 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} Picea sitchensis       11  100  NA  NA 2010 11.96667
# 3 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} Picea sitchensis       10  100  NA  NA 2010 11.96667
# 4 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} Picea sitchensis       12  100  NA  NA 2010 11.96667
# 5 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} Picea sitchensis       10  100  NA  NA 2010 11.96667
# 6 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} Picea sitchensis       13  100  NA  NA 2010 11.96667

xyTree <- TreeData[[2]]
head(xyTree)
#                                       id       y         x
# 1 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} 53.0327 -3.910179
# 2 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} 53.0327 -3.910179
# 3 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} 53.0327 -3.910179
# 4 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} 53.0327 -3.910179
# 5 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} 53.0327 -3.910179
# 6 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} 53.0327 -3.910179

plots_trees_nested_agb_sd <- sd_tree(plotTree, xyTree, 'Europe')
# The reference dataset contains 77 wood density values
# Your taxonomic table contains 1 taxa
# Using actual tree height from the provided plot data.
# Warning message:
# In BIOMASS::getWoodDensity(genus = plot$genus, species = plot$species,  :
#   DRYAD data only stored 77 wood density values in your region of interest. You could provide additional wood densities (parameter addWoodDensityData) or widen your region (region="World")

print(plots_trees_nested_agb_sd)
#                                  PLOT_ID   POINT_X  POINT_Y SIZE_HA AVG_YEAR AGB_T_HA   sdTree
# 1 {197A519B-7AF1-4374-BDD8-81FA03CCC54F} -3.910179 53.03270    0.01     2010 191.1393 27.52151
# 2 {224FD85B-6B64-4075-B51F-BA2D390AB4EF} -3.908944 53.03241    0.01     2010 243.5213 68.40316
# 3 {22E69595-D9BE-4BB1-8660-E2E7457E2068} -3.910026 53.03247    0.01     2010 302.0629 42.43342
# 4 {89B25687-EDA0-4704-B00D-93EDDA162779} -3.909820 53.03270    0.01     2010 287.5870 47.53895
# 5 {B4980FC2-4DA9-4897-93E2-07D03233D2F4} -3.908851 53.03252    0.01     2010 427.2943 90.30131

```



### Lidar reference data

Let's see an example where lidar data is used as plot data reference. In the example below we use sample lidar data provided with the package:


```{r lidar_plots, eval=FALSE}

# Sample lidar data folder location:
slb.agb.dir <- sample_lidar_folder("SustainableLandscapeBrazil_v04/SLB_AGBmaps")
slb.cv.dir <- sample_lidar_folder("SustainableLandscapeBrazil_v04/SLB_CVmaps")

# Obtaining the Coeficient of Variation for each cell in the dataset:
slb.cv <- RefLidar(slb.cv.dir)
# Enter raster type (AGB, CV, or SD): cv
# User input needed to extract PLOT_ID from filename(s)...
# File(s) used to extract data: BON_A01_2018_CV_100m.tif
# Enter numeric index of the first letter of PLOT_ID: 1
# Enter numeric index of the last letter of PLOT_ID: 7
# User input needed to extract YEAR from filename(s)...
# File(s) used to extract data: BON_A01_2018_CV_100m.tif
# Enter numeric index of the first letter of YEAR: 9
# Enter numeric index of the last letter of YEAR: 12

tail(slb.cv)
#      PLOT_ID   POINT_X   POINT_Y        CV AVG_YEAR
# 1135 BON_A01 -67.28973 -9.889289 0.3526200     2018
# 1136 BON_A01 -67.28882 -9.889289 0.3432670     2018
# 1137 BON_A01 -67.28791 -9.889289 0.3390796     2018
# 1138 BON_A01 -67.28700 -9.889289 0.3368152     2018
# 1170 BON_A01 -67.28882 -9.890197 0.3561200     2018
# 1171 BON_A01 -67.28791 -9.890197 0.3443113     2018

# Obtaining the Above Ground Biomass for each cell in the dataset:
plots_lidar <- RefLidar(slb.agb.dir)
# Enter raster type (AGB, CV, or SD): agb
# User input needed to extract PLOT_ID from filename(s)...
# File(s) used to extract data: BON_A01_2018_AGB_100m.tif
# Enter numeric index of the first letter of PLOT_ID: 1
# Enter numeric index of the last letter of PLOT_ID: 7
# User input needed to extract YEAR from filename(s)...
# File(s) used to extract data: BON_A01_2018_AGB_100m.tif
# Enter numeric index of the first letter of YEAR: 9
# Enter numeric index of the last letter of YEAR: 12

tail(plots_lidar)
#      PLOT_ID   POINT_X   POINT_Y      AGB AVG_YEAR
# 1135 BON_A01 -67.28973 -9.889289 214.9921     2018
# 1136 BON_A01 -67.28882 -9.889289 227.4593     2018
# 1137 BON_A01 -67.28791 -9.889289 233.1332     2018
# 1138 BON_A01 -67.28700 -9.889289 236.3842     2018
# 1170 BON_A01 -67.28882 -9.890197 210.6665     2018
# 1171 BON_A01 -67.28791 -9.890197 226.1070     2018

# Calculating the Standard Deviation (SD = CV * AGB) and adding it to the final plot data:
plots_lidar$sdTree <- slb.cv$CV * plots_lidar$AGB
tail(plots_lidar)
#      PLOT_ID   POINT_X   POINT_Y      AGB AVG_YEAR   sdTree
# 1135 BON_A01 -67.28973 -9.889289 214.9921     2018 75.81052
# 1136 BON_A01 -67.28882 -9.889289 227.4593     2018 78.07928
# 1137 BON_A01 -67.28791 -9.889289 233.1332     2018 79.05070
# 1138 BON_A01 -67.28700 -9.889289 236.3842     2018 79.61778
# 1170 BON_A01 -67.28882 -9.890197 210.6665     2018 75.02255
# 1171 BON_A01 -67.28791 -9.890197 226.1070     2018 77.85120

# Add biome information
plots_lidar <- BiomePair(plots_lidar)

# Add SIZE_HA column (required for calculateTotalUncertainty)
plots_lidar$SIZE_HA <- 1  # For 100m resolution (100m x 100m = 1 ha)

# Rename AGB column to AGB_T_HA to match what calculateTotalUncertainty expects
names(plots_lidar)[names(plots_lidar) == "AGB"] <- "AGB_T_HA"

# Calculate total uncertainty (measurement, sampling, and growth)
plots_lidar_unc <- calculateTotalUncertainty(plots_lidar, map_year = 2018, map_resolution = 100)
# Using existing sdTree values for tree_level data
# Calculating sampling uncertainty using Rejou-Mechain approach
# Loading sampling error data from package file
# Calculating growth uncertainty by biome
# Total uncertainty calculated for plot data of type: tree_level

head(plots_lidar_unc$data)
#     PLOT_ID AGB_T_HA AVG_YEAR   sdTree      ZONE         FAO.ecozone      GEZ SIZE_HA   POINT_X   POINT_Y
# 1 BON_A01 210.6665     2018 75.02255 S.America Tropical rainforest Tropical       1 -67.28882 -9.890197
# 2 BON_A01 226.1070     2018 77.85120 S.America Tropical rainforest Tropical       1 -67.28791 -9.890197
# 3 BON_A01 227.4593     2018 78.07928 S.America Tropical rainforest Tropical       1 -67.28882 -9.889289
# 4 BON_A01 233.1332     2018 79.05070 S.America Tropical rainforest Tropical       1 -67.28791 -9.889289
# 5 BON_A01 214.9921     2018 75.81052 S.America Tropical rainforest Tropical       1 -67.28973 -9.889289
# 6 BON_A01 195.5950     2018 72.22207 S.America Tropical rainforest Tropical       1 -67.29064 -9.888382
#   RS_HA ratio     sdSE sdGrowth  varPlot  sdTotal
# 1     1     1 16.34357        0 5895.496 76.78213
# 2     1     1 16.34357        0 6327.922 79.54824
# 3     1     1 16.34357        0 6363.486 79.77146
# 4     1     1 16.34357        0 6516.126 80.72252
# 5     1     1 16.34357        0 6014.347 77.55222
# 6     1     1 16.34357        0 5483.140 74.04823

# See the relative contribution of different uncertainty components
plots_lidar_unc$uncertainty_components
# measurement    sampling      growth 
#  0.91475934  0.08524066  0.00000000 
```




### Plot data is a GIS file

Let's see an example where the plot data is a GIS file. Users should know plot ID, plot size, inventory year beforehand for manual entry if ever missing in the GIS files' attributes. In the example below we use a sample shapefile provided with the package:


```{r shp_plots1, eval=FALSE}

plots_sf <- sf::st_read(sample_file("samp_shp.shp"))

print(plots_sf)
# Simple feature collection with 170 features and 3 fields
# Geometry type: POINT
# Dimension:     XY
# Bounding box:  xmin: 137.3756 ymin: 35.16216 xmax: 137.4079 ymax: 35.19799
# Geodetic CRS:  WGS 84
# First 10 features:
#    MEAN_Heigh        V       AGB                  geometry
# 1    20.99811 263.2125 127.70163 POINT (137.3777 35.16275)
# 2    20.92308 330.6375 177.22711 POINT (137.3774 35.16304)
# 3    18.03867 223.0688 123.08946  POINT (137.381 35.16245)
# 4    17.39841 172.4500  95.24402 POINT (137.3799 35.16216)
# 5    20.43833 249.2562 138.22925 POINT (137.3806 35.16245)
# 6    20.04576 235.3688 130.56611 POINT (137.3806 35.16363)
# 7    20.01860 266.9125 143.88520  POINT (137.377 35.16275)
# 8    21.07500 273.0625 151.47596 POINT (137.3803 35.16363)
# 9    19.81884 270.2000 148.59707 POINT (137.3785 35.16363)
# 10   22.15147 333.8500 182.89619 POINT (137.3788 35.16392)

```

The sample shapefile's SRS / CRS is already in EPSG:4326 / WGS84 so there is no need to transform it. In case it isn't we would precede the following step with `plots_sf <- st_transform(plots_sf, crs = 4326)`.


```{r shp_plots2, eval=FALSE}

plots_sf$longitude <- sf::st_coordinates(sf::st_centroid(plots_sf))[, 1]
plots_sf$latitude <- sf::st_coordinates(sf::st_centroid(plots_sf))[, 2]

shapefile_plots <- RawPlots(as.data.frame(plots_sf)) 
# Which column is your unique Plot ID? 
# 
# 1: Manual entry
# 2: MEAN_Heigh
# 3: V
# 4: AGB
# 5: geometry
# 6: longitude
# 7: latitude
# 
# Selection: 1
# Enter the numeric value for manual entry: 1
# Which column is your plot AGB? 
# 
# 1: Manual entry
# 2: MEAN_Heigh
# 3: V
# 4: AGB
# 5: geometry
# 6: longitude
# 7: latitude
# 
# Selection: 4
# Select longitude column 
# 
# 1: Manual entry
# 2: MEAN_Heigh
# 3: V
# 4: AGB
# 5: geometry
# 6: longitude
# 7: latitude
# 
# Selection: 6
# Which column is your latitude? 
# 
# 1: Manual entry
# 2: MEAN_Heigh
# 3: V
# 4: AGB
# 5: geometry
# 6: longitude
# 7: latitude
# 
# Selection: 7
# Select plot size column 
# 
# 1: Manual entry
# 2: MEAN_Heigh
# 3: V
# 4: AGB
# 5: geometry
# 6: longitude
# 7: latitude
# 
# Selection: 1
# Enter the numeric value for manual entry: 100
# Select year column 
# 
# 1: Manual entry
# 2: MEAN_Heigh
# 3: V
# 4: AGB
# 5: geometry
# 6: longitude
# 7: latitude
# 
# Selection: 1
# Enter the numeric value for manual entry: 2022

head(shapefile_plots)
#   PLOT_ID  POINT_X  POINT_Y  AGB_T_HA SIZE_HA FEZ GEZ AVG_YEAR
# 1       1 137.3777 35.16275 127.70163    0.01  NA  NA     2022
# 2       1 137.3774 35.16304 177.22711    0.01  NA  NA     2022
# 3       1 137.3810 35.16245 123.08946    0.01  NA  NA     2022
# 4       1 137.3799 35.16216  95.24402    0.01  NA  NA     2022
# 5       1 137.3806 35.16245 138.22925    0.01  NA  NA     2022
# 6       1 137.3806 35.16363 130.56611    0.01  NA  NA     2022

```



